Supplement B: Stationwise Evaluation

### 2018-10-31 pkg/disttree/inst/draft/supplement/supplement_distributions.Rnw
small improvement
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2349 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 149a8be8f8fa3bbed0d3f0b6c11a9868e3144036
changed numbering of supplements (A and B instead of 1 and 2)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2369 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 9093ae989a286d620c85bcfcb92b68ad87d3c504
small revisions: added information on supplementary material; now Appendix before references; typos
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2371 255f4249-2263-4417-b1ea-8498

\title{Supplement 1: Different Response Distributions}

@@

\Abstract{
  In the case study presented in the main manuscript ``Distributional Regression
  Forests for Probabilistic Precipitation Forecasting in Complex Terrain''
  the novel distributional regression forests are based on a Gaussian response distribution,
  left-censored at zero to accomodate both the non-negativity of precipitation and
  the point mass at zero. To assess the goodness of fit of this response distribution
  this supplement employs the same evaluations as in the main manuscript but based on
  two other distributional assumptions: A logistic distribution, left-censored at zero,
  is employed to potentially better capture heavy tails -- and a two-part hurdle model
  combining a binary model for zero vs.\ positive precipitation and a Gaussian model,
  truncated at zero, for the positive precipitation observations. The latter could
  potentially better discriminate determinants of the occurrence of precipitation
  vs.\ the amount of precipitation given that it occurs.
}

@@

The main manuscript employes a Gaussian distribution with probability
density function $f$ (PDF) and cumulative distribution function $F$ (CDF)
given by
%
\begin{align}
f_{\mathrm{Gaussian}}(Y; \mu, \sigma) &= 
\frac{1}{\sigma} \cdot \phi\left(\frac{Y-\mu}{\sigma}\right) \\
F_{\mathrm{Gaussian}}(Y; \mu, \sigma) &=
\Phi\left(\frac{Y-\mu}{\sigma}\right)
\end{align}
where $\phi$ and $\Phi$ are the PDF and the CDF of the standard normal
distribution $\mathcal{N}(0,1)$.

An alternative distribution with several similar properties but with
heavier tails is the logistic distribution. The PDF and CDF are given by
%
\begin{align}
f_{\mathrm{logistic}}(Y; \mu, \sigma) &= 
\frac{\exp{\left(-\frac{Y - \mu}{\sigma}\right)}}
{\sigma \cdot \left(1 + \exp{\left(-\frac{Y - \mu}{\sigma}\right)}\right)^2} \\
F_{\mathrm{logistic}}(Y; \mu, \sigma) &=
\frac{1}{1+ \exp{\left(-\frac{Y - \mu}{\sigma}\right)}}.
\end{align}
%
Both distributions have to be adapted in our case study to accomodate that
precipitation is non-negative with a point mass at zero (= no rain).
Two common strategies for this are to employ either \emph{left-censoring at zero}
or a \emph{hurdle} (or two-part) approach, combining a binary (zero vs.\ greater) and
a truncated (positive) part \citep[see][for an overview of both techniques]{Winkelmann:2006}.
\begin{itemize}
\item \emph{Censored distribution:}
For a chosen distribution with PDF $f$ and CDF $F$ the log-likelihood function of the corresponding 
zero-censored distribution is
%
\begin{equation}
\ell_{\mathrm{cens}}(\mu, \sigma; Y) = 
\begin{cases}
\log\left\{f(Y; \mu, \sigma) \right\}, & \text{if } Y > 0 \\[0.2cm]
\log\left\{F(0; \mu, \sigma) \right\}, & \text{if } Y = 0
\end{cases}
\end{equation}
%
as stated in Equation~2.1.1 of the main manuscript for the Gaussian special case.
For positive values of $Y$ the censored version of the PDF
equals the original non-censored version. For $Y=0$ the censored density
function takes the values of the probability mass lying left of the 
censoring point. Therefore, the parameters $\mu$ and $\sigma$ control
both aspects, the point mass and the positive part of the distribution.

@@

# boxplots with matplot
par(mfrow = c(1,4), mar = c(4,4,3,0.7))
matplot(t(s_df[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE,  main = "Distributional forest",
        xlab = "", ylab = "CRPS skill score", xlim = c(0.5, 2.5),
        ylim = c(-0.05, 0.05))
boxplot(s_df, add = TRUE, col = "transparent")
abline(h = 0, col = pal[5], lwd = 2)
#boxplot(crps_ss_dist~distribution, data = means_sel, 
#        subset = (method == "Distributional forest"), main = "Distributional forest",
#        ylim = c(-0.12,0.07), las = 2, ylab = "CRPS skill score")
#abline(h = 0, col = pal[5], lwd = 2)
par(mar = c(4,2.9,3,1.8))
matplot(t(s_g[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE, main = "Prespecified GAMLSS",
        xlab = "", ylab = "", xlim = c(0.5, 2.5),
        ylim = c(-0.05, 0.05))
boxplot(s_g, add = TRUE, col = "transparent", yaxt = 'n')
abline(h = 0, col = pal[5], lwd = 2)
#boxplot(crps_ss_dist~distribution, data = means_sel, 
#        subset = (method == "Prespecified GAMLSS"), main = "Prespecified GAMLSS",
#        ylim = c(-0.12,0.07), las = 2)
#abline(h = 0, col = pal[5], lwd = 2)
par(mar = c(4,1.8,3,2.9))
matplot(t(s_gb[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE, main = "Boosted GAMLSS",
        xlab = "", ylab = "", xlim = c(0.5, 2.5),
        ylim = c(-0.05, 0.05))
boxplot(s_gb, add = TRUE, col = "transparent", yaxt = 'n')
abline(h = 0, col = pal[5], lwd = 2)
#boxplot(crps_ss_dist~distribution, data = means_sel, 
#        subset = (method == "Boosted GAMLSS"), main = "Boosted GAMLSS",
#        ylim = c(-0.12,0.07), las = 2)
#abline(h = 0, col = pal[5], lwd = 2)
par(mar = c(4,0.7,3,4))
matplot(t(s_emos[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE,  main = "EMOS",
        xlab = "", ylab = "", xlim = c(0.5, 2.5),
        ylim = c(-0.05, 0.05))
boxplot(s_emos, add = TRUE, col = "transparent", yaxt = 'n')
abline(h = 0, col = pal[5], lwd = 2)
#boxplot(crps_ss_dist~distribution, data = means_sel, 
#        subset = (method == "EMOS"), main = "EMOS",
#        ylim = c(-0.12,0.07), las = 2)
#abline(h = 0, col = pal[5], lwd = 2)
@
\caption{\label{fig:box_crpss_gaussian} CRPS skill score (higher = better)
for hurdle Gaussian and censored logistic distributions in reference to the
censored Gaussian distribution (orange horizontal line). Gray lines correspond to the 15~stations and are aggregated in the boxplot.}
\end{figure}

\subsubsection*{FIXME: FURTHER GRAPHICS?}

\emph{The same data already shown in Figures~\ref{fig:lattice} 
and~\ref{fig:box_crpss_gaussian} can be displayed in further ways. However, these do 
not really add much and hence should probably be omitted.}

Figures~\ref{fig:box_crpss_EMOS} and \ref{fig:box_crps} compare the methods separately 
for each distribution, once based on CRPS skill score values with reference model EMOS 
and once based on CRPS values. Both confirm the conclusion that all three distributions
lead to almost the same results for all methods.

@@

\begin{figure}[t!]
\setkeys{Gin}{width=0.99\textwidth}
<<results_box_crpss_EMOS, fig=TRUE, echo=FALSE, height=3, width=7.2>>=
# crps skill score by method (reference: EMOS)
means$crps_ss_method <- means$crps
means$crps_ss_method[means$method == "Distributional forest"] <- 
  1 - means$crps[means$method == "Distributional forest"] / means$crps[means$method == "EMOS"]
means$crps_ss_method[means$method == "Prespecified GAMLSS"] <- 
  1 - means$crps[means$method == "Prespecified GAMLSS"] / means$crps[means$method == "EMOS"]
means$crps_ss_method[means$method == "Boosted GAMLSS"] <- 
  1 - means$crps[means$method == "Boosted GAMLSS"] / means$crps[means$method == "EMOS"]
means$crps_ss_method[means$method == "EMOS"] <- 
  1 - means$crps[means$method == "EMOS"] / means$crps[means$method == "EMOS"]

@@

@
\caption{\label{fig:box_crpss_EMOS}CRPS skill score values of each of 
the 15 observation stations, separated by method and distribution. 
EMOS is chosen as reference model.}
\end{figure}

\begin{figure}[t!]
\setkeys{Gin}{width=0.99\textwidth}
<<results_box_crps, fig=TRUE, echo=FALSE, height=3, width=7.2>>=

@@

@
\caption{\label{fig:box_crps}CRPS values of each of the 15 observation stations, 
separated by method and distribution.}
\end{figure}

>>>>

\title{Supplement A: Different Response Distributions}

@@

\Abstract{
  In the case study presented in the main manuscript ``Distributional Regression
  Forests for Probabilistic Precipitation Forecasting in Complex Terrain''
  the novel distributional regression forests are based on a Gaussian response distribution,
  left-censored at zero to accommodate both the non-negativity of precipitation and
  the point mass at zero. To assess the goodness of fit of this response distribution
  this supplement employs the same evaluations as in the main manuscript but based on
  two other distributional assumptions: A logistic distribution, left-censored at zero,
  is employed to potentially better capture heavy tails -- and a two-part hurdle model
  combining a binary model for zero vs.\ positive precipitation and a Gaussian model,
  truncated at zero, for the positive precipitation observations. The latter could
  potentially better discriminate determinants of the occurrence of precipitation
  vs.\ the amount of precipitation given that it occurs.
}

@@

The main manuscript employs a Gaussian distribution with probability
density function $f$ (PDF) and cumulative distribution function $F$ (CDF)
given by
%
\begin{align}
f_{\mathrm{Gaussian}}(Y; \mu, \sigma) &= 
\frac{1}{\sigma} \cdot \phi\left(\frac{Y-\mu}{\sigma}\right) \\
F_{\mathrm{Gaussian}}(Y; \mu, \sigma) &=
\Phi\left(\frac{Y-\mu}{\sigma}\right)
\end{align}
where $\phi$ and $\Phi$ are the PDF and the CDF of the standard normal
distribution $\mathcal{N}(0,1)$.

An alternative distribution with several similar properties but with
heavier tails is the logistic distribution. The PDF and CDF are given by
%
\begin{align}
f_{\mathrm{logistic}}(Y; \mu, \sigma) &= 
\frac{\exp{\left(-\frac{Y - \mu}{\sigma}\right)}}
{\sigma \cdot \left(1 + \exp{\left(-\frac{Y - \mu}{\sigma}\right)}\right)^2} \\
F_{\mathrm{logistic}}(Y; \mu, \sigma) &=
\frac{1}{1+ \exp{\left(-\frac{Y - \mu}{\sigma}\right)}}.
\end{align}
%
Both distributions have to be adapted in our case study to accommodate that
precipitation is non-negative with a point mass at zero (= no rain).
Two common strategies for this are to employ either \emph{left-censoring at zero}
or a \emph{hurdle} (or two-part) approach, combining a binary (zero vs.\ greater) and
a truncated (positive) part \citep[see][for an overview of both techniques]{Winkelmann:2006}.
\begin{itemize}
\item \emph{Censored distribution:}
For a chosen distribution with PDF $f$ and CDF $F$ the log-likelihood function of the corresponding 
zero-censored distribution is
%
\begin{equation}
\ell_{\mathrm{cens}}(\mu, \sigma; Y) = 
\begin{cases}
\log\left\{f(Y; \mu, \sigma) \right\}, & \text{if } Y > 0 \\[0.2cm]
\log\left\{F(0; \mu, \sigma) \right\}, & \text{if } Y = 0
\end{cases}
\end{equation}
%
as stated in Equation~2.1.1 of the main manuscript for the Gaussian special case.
For positive values of $Y$ the censored version of the PDF
equals the original non-censored version. For $Y=0$ the censored density
function takes the values of the probability mass lying left of the 
censoring point. Therefore, the parameters $\mu$ and $\sigma$ control
both aspects, the point mass and the positive part of the distribution.

@@

# boxplots with matplot
par(mfrow = c(1,4), mar = c(4,4,3,0.7))
matplot(t(s_df[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE,  main = "Distributional forest",
        xlab = "", ylab = "CRPS skill score", xlim = c(0.5, 2.5),
        ylim = c(-0.05, 0.05))
boxplot(s_df, add = TRUE, col = "transparent")
abline(h = 0, col = pal[5], lwd = 2)
#boxplot(crps_ss_dist~distribution, data = means_sel, 
#        subset = (method == "Distributional forest"), main = "Distributional forest",
#        ylim = c(-0.12,0.07), las = 2, ylab = "CRPS skill score")
#abline(h = 0, col = pal[5], lwd = 2)
par(mar = c(4,2.9,3,1.8))
matplot(t(s_g[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE, main = "Prespecified GAMLSS",
        xlab = "", ylab = "", xlim = c(0.5, 2.5),
        ylim = c(-0.05, 0.05))
boxplot(s_g, add = TRUE, col = "transparent", yaxt = 'n')
abline(h = 0, col = pal[5], lwd = 2)
#boxplot(crps_ss_dist~distribution, data = means_sel, 
#        subset = (method == "Prespecified GAMLSS"), main = "Prespecified GAMLSS",
#        ylim = c(-0.12,0.07), las = 2)
#abline(h = 0, col = pal[5], lwd = 2)
par(mar = c(4,1.8,3,2.9))
matplot(t(s_gb[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE, main = "Boosted GAMLSS",
        xlab = "", ylab = "", xlim = c(0.5, 2.5),
        ylim = c(-0.05, 0.05))
boxplot(s_gb, add = TRUE, col = "transparent", yaxt = 'n')
abline(h = 0, col = pal[5], lwd = 2)
#boxplot(crps_ss_dist~distribution, data = means_sel, 
#        subset = (method == "Boosted GAMLSS"), main = "Boosted GAMLSS",
#        ylim = c(-0.12,0.07), las = 2)
#abline(h = 0, col = pal[5], lwd = 2)
par(mar = c(4,0.7,3,4))
matplot(t(s_emos[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE,  main = "EMOS",
        xlab = "", ylab = "", xlim = c(0.5, 2.5),
        ylim = c(-0.05, 0.05))
boxplot(s_emos, add = TRUE, col = "transparent", yaxt = 'n')
abline(h = 0, col = pal[5], lwd = 2)
#boxplot(crps_ss_dist~distribution, data = means_sel, 
#        subset = (method == "EMOS"), main = "EMOS",
#        ylim = c(-0.12,0.07), las = 2)
#abline(h = 0, col = pal[5], lwd = 2)
@
\caption{\label{fig:box_crpss_gaussian} CRPS skill score (higher = better)
for hurdle Gaussian and censored logistic distributions in reference to the
censored Gaussian distribution (orange horizontal line). Gray lines
correspond to the 15~stations that are aggregated in the boxplot.}
\end{figure}

%\subsubsection*{FIXME: FURTHER GRAPHICS?}

%\emph{The same data already shown in Figures~\ref{fig:lattice} 
%and~\ref{fig:box_crpss_gaussian} can be displayed in further ways. However, these do 
%not really add much and hence should probably be omitted.}

%Figures~\ref{fig:box_crpss_EMOS} and \ref{fig:box_crps} compare the methods separately 
%for each distribution, once based on CRPS skill score values with reference model EMOS 
%and once based on CRPS values. Both confirm the conclusion that all three distributions
%lead to almost the same results for all methods.

@@

%\begin{figure}[t!]
%\setkeys{Gin}{width=0.99\textwidth}
<<results_box_crpss_EMOS, fig=TRUE, echo=FALSE, height=3, width=7.2, eval=FALSE>>=
# crps skill score by method (reference: EMOS)
means$crps_ss_method <- means$crps
means$crps_ss_method[means$method == "Distributional forest"] <- 
  1 - means$crps[means$method == "Distributional forest"] / means$crps[means$method == "EMOS"]
means$crps_ss_method[means$method == "Prespecified GAMLSS"] <- 
  1 - means$crps[means$method == "Prespecified GAMLSS"] / means$crps[means$method == "EMOS"]
means$crps_ss_method[means$method == "Boosted GAMLSS"] <- 
  1 - means$crps[means$method == "Boosted GAMLSS"] / means$crps[means$method == "EMOS"]
means$crps_ss_method[means$method == "EMOS"] <- 
  1 - means$crps[means$method == "EMOS"] / means$crps[means$method == "EMOS"]

@@

@
%\caption{\label{fig:box_crpss_EMOS}CRPS skill score values of each of 
%the 15 observation stations, separated by method and distribution. 
%EMOS is chosen as reference model.}
%\end{figure}

%\begin{figure}[t!]
%\setkeys{Gin}{width=0.99\textwidth}
<<results_box_crps, fig=TRUE, echo=FALSE, height=3, width=7.2, eval=FALSE>>=

@@

@
%\caption{\label{fig:box_crps}CRPS values of each of the 15 observation stations, 
%separated by method and distribution.}
%\end{figure}

### 2018-10-31 pkg/disttree/inst/draft/supplement/supplement_stationwise.Rnw
small revisions: added information on supplementary material; now Appendix before references; typos
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2371 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 1c9e1be81d8df61a522af0f5397b6a7b825a7557

The Resiudal QQ plots and the PIT histograms visualize the same values,
however, on different scales. For each observation $y$ of the test sample 
the corresponding PIT value is calculated by applying the probability function 
$F(y; \hat{\mu}, \hat{\sigma})$
with the estimated distribution parameters plugged in. The resulting values
can either be plotted in a histogram (PIT histogram) or be plotted against
the corresponding quantiles of a standard normal distribution (residual QQ plot)
for a continuous variable $Y$. To account for the point mass at 0, for each 
observation equal to 0 the corresponding PIT value is multiplied with a randomized
value following a uniform distribution on \[0,1\]. In that way these values
are apportioned on those bins of the histogram that would not contain any
values because no observation is beneath 0.

>>>>

The Resiudal QQ plots and the PIT histograms visualize the same values,
however, on different scales. For each observation $y$ of the test sample 
the corresponding PIT value is calculated by applying the probability function 
$F(y; \hat{\mu}, \hat{\sigma})$
with the estimated distribution parameters plugged in. The resulting values
can either be plotted in a histogram (PIT histogram) or be plotted against
the corresponding quantiles of a standard normal distribution (residual QQ plot)
for a continuous variable $Y$. To account for the point mass at 0, for each 
observation equal to 0 the corresponding PIT value is multiplied with a randomized
value following a uniform distribution on $[0,1]$. In that way these values
are apportioned on those bins of the histogram that would not contain any
values because no observation is beneath 0.
